Dependently Typed Languages in Statix

### 2023-04-19 pres-thesis/presentation.tex
Start on thesis presentation
Work on thesis defense presentation

\documentclass[aspectratio=43]{beamer}
\usepackage{csquotes}
\usepackage{calc}
\usepackage[absolute,overlay]{textpos}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{mathtools}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{comment}
\usepackage{siunitx}
\usepackage{MnSymbol,wasysym}
\usepackage{array}
\usepackage{qrcode}

@@

\begin{frame}{My Thesis}
	\begin{itemize}
		\item Followed Compiler Construction A+B from Eelco
		\item Open question from Eelco: \textbf{How powerful is Statix?}
	\end{itemize}
\end{frame}

\begin{frame}[fragile]{Background: What are Dependent Types?}
\begin{itemize}
	\item Types may depend on values!
	\begin{exampleblock}{Example}
		\texttt{concat : (A: Set) -> (n m : Nat) -> Vec A n -> Vec A m
			\\ \hspace*{48pt} -> Vec A (n + m)}
	\end{exampleblock}
% - n occurs insie a type `Vec`!
% - The type system enforces these lengths are always correct
% - Challenge: Type checking may require evaluating arbitrary terms to decide equality
	\item Proof assistants
	\item For example: Agda, Coq, Lean, ...
% Can be used to write proofs about code
\end{itemize}

@@

@@

	\begin{exampleblock}{Example 1}
		\texttt{($\backslash$v: Type. v) Type}
	\end{exampleblock}

	\begin{exampleblock}{Example 2}
		\texttt{let f = $\backslash$T: Type. $\backslash$x: T. x; \\
f (T: Type -> Type) ($\backslash$y: Type. y)
		}
	\end{exampleblock}

@@

\begin{frame}[fragile]{Type Checking}
	\begin{block}{Type checking relation}
		\texttt{typeOfExpr : scope * Expr -> Expr}
	\end{block}
	% Returns an Expr, not a Type
	% Scope -> Two different types of nodes
	\begin{block}{How do we use scopes?}
		A scope is used as a replacement for an environment and a context.\\
		One edge \verb|p|. One relation \texttt{name $\rightarrow$ NameEntry}, NameEntry is either:
		\begin{itemize}
			\item NType: Stores a type (Corresponds to a context)
			\item NSubst: Stores a substitution (Corresponds to an environment)
		\end{itemize}
	\end{block}
\end{frame}

@@

>>>>

\documentclass[aspectratio=43]{beamer}
\usepackage{csquotes}
\usepackage{calc}
\usepackage[absolute,overlay]{textpos}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{mathtools}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{comment}
\usepackage{siunitx}
\usepackage{MnSymbol,wasysym}
\usepackage{array}
\usepackage{qrcode}
\usepackage{listings}

@@

\begin{frame}[fragile]{Background: What is Spoofax?}
	A language designer's workbench with everything you need to design a programming language.
	\begin{itemize}
		\item Declarative, using meta-languages
	\end{itemize}
\end{frame}

\begin{frame}[fragile]{Background: What is Spoofax?}
SDF3, Statix and Stratego (and more)

\begin{figure}
	\includegraphics[width=0.32\linewidth]{img/semantics}
	\includegraphics[width=0.32\linewidth]{img/syntax}
	\includegraphics[width=0.32\linewidth]{img/transformation}
\end{figure}
\end{frame}

\begin{frame}[fragile]{Background: What is Spoofax?}

\begin{exampleblock}{Example}
\begin{lstlisting}
int x = 5;
int y = x + 3;
return x + y;
\end{lstlisting}
\end{exampleblock}

\begin{figure}
	\includegraphics[width=0.7\linewidth]{img/screenshot005}
\end{figure}

% The logic I just explained would be written down in Statix
% Like: When you encounter a let, create a new scope...

@@

\begin{frame}[fragile]{Background: What are Dependent Types?}
\begin{itemize}
	\item Types may depend on values!
	\begin{exampleblock}{Example}
		\texttt{concat : (A: Set) -> (n m : Nat) -> Vec A n -> Vec A m
			\\ \hspace*{48pt} -> Vec A (n + m)}
	\end{exampleblock}
% - n occurs insie a type `Vec`!
% - The type system enforces these lengths are always correct
% - Challenge: Type checking may require evaluating arbitrary terms to decide equality
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Background: What are Dependent Types?}
	Type checking requires evaluation
	\begin{exampleblock}{Example 1}
		\texttt{let T = if false then Int else Bool end;\\let b: T = true;}
	\end{exampleblock}
\end{frame}

\begin{frame}[fragile]{Background: What are Dependent Types?}
	Why are dependent types useful?
	\begin{itemize}
		\item Proof Assistants: Agda, Lean, Coq, etc
	\end{itemize}

	% We're looking only at signatures
	\begin{exampleblock}{Sorted lists}
\texttt{sort : List t -> List t\\sort\_sorted : (v : List t) -> IsSorted (sort v)}
	\end{exampleblock}
\end{frame}

@@

	\begin{block}{Syntax Definition}
		\begin{figure}
			\includegraphics[width=0.7\linewidth]{img/screenshot006}
		\end{figure}
	\end{block}

	\begin{exampleblock}{Example}
		\texttt{let f = $\backslash$T: Type. $\backslash$x: T. x; \\
f (T: Type -> Type) ($\backslash$y: Type. y)
		}
	\end{exampleblock}

@@

\begin{frame}[fragile]{Type Checking}
	\begin{figure}
		\includegraphics[width=0.9\linewidth]{img/screenshot007}
	\end{figure}
	% We define the language using inference rules
	% How can we express those rules in Statix?
\end{frame}

@@

\begin{frame}[fragile]{Type Checking}
	\begin{block}{How do we use scopes?}
		A scope is used as a replacement for an environment and a context.\\
		One edge \verb|p|. One relation \texttt{name $\rightarrow$ NameEntry}, NameEntry is either:
		\begin{itemize}
			\item NType: Stores a type (Corresponds to a context)
			\item NSubst: Stores a substitution (Corresponds to an environment)
		\end{itemize}
	\end{block}
\end{frame}
